An Incomplete Overview of some Applications of Game Theory to Patient Flow

### 2014-05-12 paper.tex
added a lot of stuff

\usepackage{tikz}
\usepackage{parskip}

@@

Talk about idea.
Talk about Markov model.
Talk about proof that pure strategy NE exists.
Show how can be used to influence t.

>>>>

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}

\usepackage[parfill]{parskip}

\newcommand{\NH}{\text{NH}}
\newcommand{\RG}{\text{RG}}

@@

The main proposition of the work requested by the ABUHB was to develop a mathematical model of bed occupancies at the CCUs at RG and NH.
After an initial analysis of the data, behavioural aspects became apparent; for example, delaying patients discharge if there was no pressure on CCU beds, or admitting fewer patients if bed occupancy levels were high.
As a result of this, a state-dependent queueing model has been developed, which includes the dependency of admission rate on actual occupancy \cite{Komenda2013}.
This state dependent model was applied to both NH and RG separately.
It is however obvious that the actions of one CCU impact on the other CCU, as diversion of patients from one CCU to the other sometimes occurs.
A pictorial representation of the situation is given in Figure \ref{diagramofdiversion}.

\begin{figure}[!htbp]
\begin{center}
\begin{tikzpicture}
\node (A) at (0,0) [draw] {RG};
\node (B) at (2,0) [draw] {NH};
\draw [->] (-1,0) -- (A);
\draw [->] (3,0) -- (B);
\draw [->] (-1,0) -- ($(A)+(-.5,0)$) -- ($(A)+(-.5,.5)$) -- ($(A)+(.5,.5)$) -- ($(A)+(.5,.1)$) -- ($(B)+(-.4,.1)$);
\draw [->] (3,0) -- ($(B)+(.5,0)$) -- ($(B)+(.5,-.5)$) -- ($(B)+(-.5,-.5)$) -- ($(B)+(-.5,-.1)$) -- ($(A)+(.4,-.1)$);
\draw [->] (3,0) -- (B);
\node at ($(A) + (0,.8)$) {Divert?};
\node at ($(B) + (0,-.8)$) {Divert?};
\end{tikzpicture}
\caption{Diagrammatic representation of CCU interaction through patient diversion.}\label{diagramofdiversion}
\end{center}
\end{figure}

The capacity thresholds are denoted as $K_{h}\in\mathbb{Z}$ for $h\in\{\NH,\RG\}$. Note that $0\leq K_h\leq c_h$.

\begin{figure}[!htbp]
\begin{center}
    \begin{tikzpicture}
        \draw  (0,0) rectangle (4,3);
        \draw [dashed, red, thick] (0,1) -- (4,1);
        \draw [dashed, red, thick] (3,0) -- (3,3);
        \node at (-.5,1) {$K_{\NH}$};
        \node at (3,3.25) {$K_{\RG}$};
        \node at (1.5,2) {$\lambda_{h}^{(a)}$};
        \node at (3.5,2) {$\lambda_{h}^{(b)}$};
        \node at (1.5,.5) {$\lambda_{h}^{(c)}$};
        \node at (3.5,.5) {$\lambda_{h}^{(d)}$};
    \end{tikzpicture}
\caption{General arrival rates for each CCU at each region, where $h\in\{\NH, \RG\}$} \label{arrivalrateregions}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\input{Images/markovdiagram.tex}
\caption{Generic Markov chain underpinning the queueing model of this paper} \label{mc}
\end{center}
\end{figure}

Utilities will be of interest when this queueing theoretical model will be inserted in the game theoretical model.
Throughput of patients is a natural choice of utility given that most hospitals are financially rewarded per served patient \cite{Pate2009}.
For each threshold pair $(K_{\NH},K_{\RG})$, the utilisation rate $U_h$ and throughput $T_h$ can easily be obtained for each CCU: $h\in\{\text{NH},\text{RG}\}$, using the following formulas:

$$U_{h}={{\sum_{n=0}^{c_h} nP^{(h)}(n)}\over{c_{h}}}$$
$$T_{h}=\mu_h \sum _{n=0}^{c_h} nP^{(h)}(n)$$

where $P^{(h)}=P^{(h)}(K_{\NH},K_{\RG})$ is the steady state probability distribution function (obtained from the corresponding transition matrix $Q=Q(K_{\NH},K_{\RG})$) for $h\in\{\text{NH},\text{RG}\}$.

For $c_{\NH}=8$, $c_{\RG}=16$, $\lambda_{\NH}=(\lambda_{\NH}^{(a)},\lambda_{\NH}^{(b)},\lambda_{\NH}^{(c)},\lambda_{\NH}^{(d)})=(1.5,3.74,0,0)$, $\lambda_{\RG}=(\lambda_{\RG}^{(a)},\lambda_{\RG}^{(b)},\lambda_{\RG}^{(c)},\lambda_{\RG}^{(d)})=(2.24,0,3.74,0)$ and $(K_{\NH},K_{\RG})=(6,12)$, the steady state probabilities for each CCU are given in Figure \ref{steady_state_probs}.

\begin{figure}[!htbp]
\hspace{2cm}For all $h\in\{\text{NH}, \text{RG}\}$ minimise:
$$\left(U_{h}-t\right)^2$$
\hspace{2cm}Subject to:
\begin{align}
0\leq & K_h \leq c_{h}\nonumber\\
&K_h \in  \mathbb{Z}\nonumber
\end{align}
\caption{The optimisation problem underlying the game}\label{model1}
\end{figure}

This game is equivalent to a bi matrix game with restriction to pure strategies where both players aim to get their utilisation as close as possible to a certain target.
As such a Nash Equilibrium is not guaranteed by traditional game theoretical results \cite{Nash1950}, but based on discussions with ABUHB, long term threshold policies are a realistic consideration.

The following result is a sufficient condition for the existence of an equilibrium:\\

\textbf{Theorem.}

Let $f_{h}(k):[1,c_{\bar h}]\to[1,c_h]$ be the best response of player $h\in\{\NH, \RG\}$ to the diversion threshold of $\bar h\ne h$ ($\bar h\in\{\NH, \RG\}$).

If $f_{h}(k)$ is a non-decreasing function in $k$ then the game of Figure \ref{model1} has at least one Nash Equilibrium.

\textbf{proof}

The function $f_h$ is well defined as it maximises a continuous function over a finite discrete set (in case of multiple values that minimize $U_h$, it is assumed that $f_h$ returns the lowest such value).

As such if $f_h$ is non-decreasing then it is in fact a stepwise non-decreasing function.
If we consider $f_{\NH}$ and $f_{\RG}$ plotted on the same axis (so that the domain of $f_{\NH}$ is the $x$-axis and the domain of $f_{\RG}$ is the y-axis) it is obvious to see that the functions must intersect at some point as shown in Figure \ref{exampleforproof}.

\begin{figure}[!htbp]
$$\begin{array}{cc}
\includegraphics[width=.5\textwidth]{./Images/best_responses_ex_for_proof_NH.pdf}&
\includegraphics[width=.5\textwidth]{./Images/best_responses_ex_for_proof_RG.pdf}
\end{array}$$
\caption{Plots of $f_{\NH}(K_{\RG})$ and $f_{\RG}(K_{\NH})$}\label{exampleforproof}
\end{figure}

This point of intersection corresponds to a Nash Equilibrium of Figure \ref{model1}.
$\square$

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=12cm]{./Images/argminPoAmodel2.pdf}
\caption{Lowest value of $t$ ensuring PoA$=1$} \label{mintargetvdemandmodel2}
\end{center}
\end{figure}
